---
title: "RChavez_Livesession3assignment"
author: "Ruben Chavez"
date: "November 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 1

The following code is Git and completes the described actions:  
Note: The code assumes you are in a pre-existing directory in which you wish to create your new directory.

* Create a new directory

```{r Create a new directory with Git}
#mkdir Session-3-Homework
```

* Navigate to the appropriate directory

```{r Navigate to the appropriate directory with Git}
#cd Session-3-Homework
```

* Clone the repository, "https://github.com/caesar0301/awesome-public-datasets" to the directory

```{r clone the repository to the new directory with Git}
#git clone [https://github.com/caesar0301/awesome-public-datasets]
```

#Question 2
##a.Read in the titanic.csv file

```{r read in file}
df <- read.csv("~/GitHub/Session-3-Homework/awesome-public-datasets/Datasets/titanic.csv/titanic.csv")
```

##b.Outbut the respective count of females and Males and Plot the frequency of females and Males

```{r count of females and males}
table(df$sex)
```


```{r Plot of Frequency of Females and Males}
barplot.default(table(df$Sex), main = "Frequency by Gender")
```

##c Use an apply funtion to outpout the means of Age, Fare, and Survival

```{r sapply to return mean of columns for age fare and survival}
sapply(df[, c("Age", "Fare", "Survived")], mean, na.rm=TRUE)
```


#Question 3

```{r read sleep_data_01}
sleep_data_01 <- read.csv("http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv")
```



```{r helper function}
HelperFunction<- function(x){
  MedianAge <- median(x[,"Age"], na.rm = TRUE)
  MaxSleepDuration <- max(x[,"Duration"], na.rm = TRUE)
  MinSleepDuration <- min(x[,"Duration"], na.rm = TRUE)
  MeanRSES <- mean(x[,"RSES"], na.rm = TRUE)
  SDRSES <- sd(x[,"RSES"], na.rm = TRUE)
  Report <- data.frame("Median Age" =  MedianAge, "Mean RSES" =MeanRSES/5,
                       "SD RSES" = SDRSES/5,"Range of Duration" = 
                         MaxSleepDuration - MinSleepDuration)

  names(Report) <- c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
  Report <- round(Report, digits =2)
  return(Report)
}

HelperFunction(sleep_data_01)
```

#Question 4

##a.
```{r install fivethirtyeight}
#install.packages("fivethirtyeight")
```

##b.
```{r assign 22nd data set to df}
df <- get(data(list=data(package="fivethirtyeight")$results[22,"Item"],package = "fivethirtyeight"))
```

##c.

```{r url for related new story}
#http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

##d.

```{r dimensions and column names for df}
dim(df)
names(df)
```

#Question 5

##a

```{r Column names and count}
names(df)
ncol(df)
```

##b

```{r count of unique values in major_category}
major_count <- as.data.frame(table(df$major_category))
```

##c

```{r barplot of major_count}
par(las=1)
par(mar=c(5,17,4,2))
barplot.default(height = major_count$Freq,names.arg = major_count$Var1, horiz = TRUE, col = "aquamarine2", xlab = "Count of Major Category",main = "Frequency of Unique Major Categories")
title(ylab = "Major Categories", mgp=c(15,1,0))
```

##d

```{r write df to a csv file}
write.csv(df, file = "College Recent Graduates", row.names = FALSE)
```

#Question 6

https://github.com/rechavez/Ruben-Chavez-SMU-MSDS-Homework
